### 2016-02-16 mcmc.tex
Gelman--Rubin

HOGG: Introduce and explain the Gelman--Rubin
diagnostic\note{\citet{gelmanrubin}.}, which is also a good idea, and
which tests something slightly different than the autocorrelation
time.

>>>>

One simple and sensible test of convergence is the Gelman--Rubin
diagnostic \note{\citet{gelmanrubin}.}, which compares the variance
(in one parameter, or your most important parameter, or all
parameters) \emph{within} a chain to the variance \emph{across}
chains.
This requires running multiple chains, and looking at the empirical
variance of the parameter away from its mean within each individual
chain, and comparing it to the variance in the mean of that parameter
across chains, inflated to be a mean per-sample variance.
What Gelman and Rubin do with these variances specifically is
sensible, but the important question of convergence is whether, as the
chains get longer, these two variances asymptotically reach stable
values, and that those two values agree.
The Gelman--Rubin diagnostic is related to the autocorrelation time,
in that it will only deliver success when the chains are much longer
than an autocorrelation time.
It can also be used to assess the accuracy of the chain for performing
integrals---by replacing the mean of the parameter with whatever
integral is of most interest.
